---
title: "PML Project"
author: "tameemmd"
date: "Sunday, March 22, 2015"
output: html_document
---

#Executive Summary

In this report we'll develop an algorithm to classify the way an exercise is made by using accelerometer variables as predictors. We develop one model based in knn . The accuracy is 96%.


##Reading data and loading libraries

First we'll read the test and training data. We have to make a validation partition from the training data.

```{r}
library(caret)
traindata<-read.csv("pml-training.csv")
inTrain <- createDataPartition( traindata$classe, p=0.65, list=FALSE);
training <- traindata[ inTrain,];
validation <- traindata[ -inTrain,];
testdata<-read.csv("pml-testing.csv")
```

Now, we'll look at the data we have read. As we don't know how many predictors we have, we'll first make a count of columns.

```{r}
length(names(traindata))
```


We notice a high number of predictors, so we could use a way to reduce them.


#Cleaning data

An out-of-the-hat way is to remove all the columns that have NAs and blank strings.

```{r}
training2 <- training[,  !apply( is.na(training), 2,any)]
training2 <- training2[, !apply( training2=="",   2,any)]
length(names(training2))
```

#Models creation & Accuracy

We have now 60 predictors, so we'll try to use them to fit a model. We'll start using the knn method, to determine the closest neighbors of the each prediction classes. We'll make a preprocessing with pca.

```{r}
model1<-train(classe~.,data=training2, method = "knn",preProcess="pca")
```

We'll now use our model to predict the classes of our test data, to calculate the
sample error.

```{r}
confusionMatrix( predict( model1, validation ), validation$classe );
```


#Conclusion

We have demonstrated that the knn method works well for this classification problem.We achieve about 96% accuracy in our prediction. While it is possible to build other types of models, targeting to improve accuracy beyond the point reached here could be a mistake resulting in over fitting the data.